---
title: "Machine Learning Final Project"
author: "Jordan Mantha"
date: "8/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(dplyr)
library(parallel)
library(doParallel)
set.seed(4321)
```

## Loading and Preprocessing

The training data and the data used for the testing quiz are loaded in. Since there are many columns that have mostly missing data, columns with greater than 75% NAs are removed from both data sets. Columns that are not relevant (timestamps, etc.) are also removed. We are left with 52 predictors in addition to the classe outcome column.

```{r cache=TRUE}
trainingData <- read.csv("pml-training.csv", na.strings = c("NA", ""))
quiz <- read.csv("pml-testing.csv")

bad <- c()
for(i in names(trainingData)) {
        nas <- sum(is.na(trainingData[[i]]))
        if ( nas != 0) {
                bad <- c(bad, i)
        }
}
trainingData <- trainingData %>% select(-(X:num_window),-bad)
quiz <- quiz %>% select(-(X:num_window),-bad)

inTrain <- createDataPartition(trainingData$classe, p = 0.7, list = FALSE)
training <- trainingData[inTrain, ]
testing <- trainingData[-inTrain, ]
dim(training)
dim(testing)
```

## Training

The random forest method is used with a 5-fold "cv" cross-validation method.

```{r cache=TRUE}
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
trControl <- trainControl(method = "cv", number = 5, allowParallel = TRUE)
modFit <- train(classe ~ ., method = "rf", trControl = trControl, data = training)
stopCluster(cluster)
registerDoSEQ()
```

## Results

The results from the model fit are given below:
```{r echo=TRUE}
print(modFit)
plot(modFit)
```
The out-of-bag estimate of error is 0.75%

To assess the accuracy we look at the confusion matrix on predictions of the testing portion of the training data with the real classe variable in the testing data.

```{r}
modFitPred <- predict(modFit, training)
modFitCM <- confusionMatrix(modFitPred,training$classe)
print(modFitCM)
ac <- modFitCM$overall["Accuracy"]
```

The accuracy of the prediction is `r paste0(round(ac * 100, 2), "%")`.

Lastly, we predict the values of the testing data for the quiz:
```{r}
predict(modFit, quiz)
```